---
title: "main"
output: html_document
date: "2026-05-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The `r BiocStyle::Biocpkg("SpatialData")` package provides an R interface to the 
[SpatialData](https://spatialdata.scverse.org) framework, a unified ecosystem 
for handling spatial omics data. Developed as part of the 
[scverse](https://scverse.org) project [@Virshup2023-scverse], `SpatialData` aims to solve the 
challenges of integrating diverse spatial datasets—including imaging, spatial 
transcriptomics, and proteomics—by employing the 
[OME-NGFF (Next Generation File Format)](https://ngff.openmicroscopy.org) 
standard [@Marconato2025-SpatialData].

The Python implementation and core specifications can be found at the 
[official SpatialData website](https://spatialdata.scverse.org).

## Data Representation in R

The core of the package is the `SpatialData` class, which serves as a container 
for multiple layers of spatial information:

- **Images**: Multiscale, multi-channel raster data, represented as `ZarrArray` objects via the `r BiocStyle::Biocpkg("Rarr")` and `r BiocStyle::Biocpkg("ZarrArray")` packages.
- **Labels**: Segmentation masks (e.g., cell or nuclei boundaries), also stored as `ZarrArray`s.
- **Points**: High-resolution coordinates (e.g., individual transcript molecules), handled as `r BiocStyle::CRANpkg("arrow")` tables or `r BiocStyle::CRANpkg("duckdb")`-backed data frames for efficient out-of-memory processing.
- **Shapes**: Polygonal regions of interest or circular markers, represented similarly to points but with geometric metadata.
- **Tables**: Functional annotations (e.g., gene expression counts), stored as `r BiocStyle::Biocpkg("SingleCellExperiment")` objects.

# Getting Started

First, we load the necessary libraries:

```{r load-libs}
library(SpatialData)
library(SingleCellExperiment)
```

## Reading Data

`SpatialData` objects are typically read from `.zarr` stores. The package provides 
the `readSpatialData()` function to ingest an entire store. For this demonstration, 
we use a toy dataset included in the package.

```{r read-data}
# Path to a toy dataset included in the package
path <- system.file("extdata", "blobs.zarr", package="SpatialData")

# Read the complete SpatialData object
# sd <- readSpatialData(path)
# sd
tm <- SpatialData::readImage(file.path(path, "images/blobs_image"))
tm
data(tm)
```